---
title: "MutualInformation"
author: "Brody Kendall"
date: "5/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(igraph)
library(infotheo)
library(ggplot2)
```

```{r}
dat_seq = readRDS("~/Documents/Northwestern/DNA_Cyclizability/data/Created/processed_seq.rds")
dat = readRDS("~/Documents/Northwestern/DNA_Cyclizability/data/Created/processed.rds")
```


```{r}
cutoffs_0_25 = quantile(dat_seq$C0, c(0, 0.25)) 
cutoffs_25_50 = quantile(dat_seq$C0, c(0.25, 0.5))
cutoffs_50_75 = quantile(dat_seq$C0, c(0.5, 0.75))
cutoffs_75_100 = quantile(dat_seq$C0, c(0.75, 1))
```

```{r}
dat_q1 = dat_seq %>%
  filter(C0 >= cutoffs_0_25[1] & C0 <= cutoffs_0_25[2]) %>%
  select(all_of(ps1), all_of(ps2), C0)

d_q1 = as.data.frame(matrix(nrow = 1225, ncol = 3))
colnames(d_q1) = c("Var1", "Var2", "MI")
z=1
for(i in 1:49) {
  for(j in (i+1):50) {
    d_q1[z, "Var1"] = i
    d_q1[z, "Var2"] = j
    d_q1[z, "MI"] = mutinformation(dat_q1[,paste0("X", i, "mono")], dat_q1[,paste0("X", j, "mono")])
    z=z+1
  }
}

d_q1.ord <- d_q1

vertex.size = 5
edge.scalar = 200
text.size = 10
number.vertices = 150
d_q1 = d_q1 %>%
  arrange(desc(MI))
g_q1 = graph_from_data_frame(d_q1[1:number.vertices,], directed=FALSE)
# plot(g_q1, vertex.size = vertex.size, edge.width = E(g_q1)$MI*edge.scalar, cex = text.size, layout = layout.circle, order = order(as.numeric(names(V(g_q1)))))
# plot(g_q1, vertex.size = vertex.size, edge.width = E(g_q1)$MI*edge.scalar, cex = text.size, layout = layout.circle)
coords_q1 <- layout_in_circle(g_q1, order =
                                order(as.numeric(names(V(g_q1)))))
plot(g_q1, vertex.size = vertex.size, edge.width = E(g_q1)$MI*edge.scalar, cex = text.size, layout = coords_q1, vertex.shape = "none")
```

```{r}
dat_q2 = dat_seq %>%
  filter(C0 >= cutoffs_25_50[1] & C0 <= cutoffs_25_50[2]) %>%
  select(all_of(ps1), all_of(ps2), C0)

d_q2 = as.data.frame(matrix(nrow = 1225, ncol = 3))
colnames(d_q2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:49) {
  for(j in (i+1):50) {
    d_q2[z, "Var1"] = i
    d_q2[z, "Var2"] = j
    d_q2[z, "MI"] = mutinformation(dat_q2[,paste0("X", i, "mono")], dat_q2[,paste0("X", j, "mono")])
    z=z+1
  }
}

d_q2 = d_q2 %>%
  arrange(desc(MI))
g_q2 = graph_from_data_frame(d_q2[1:number.vertices,], directed=FALSE)

coords_q2 <- layout_in_circle(g_q2, order =
                                order(as.numeric(names(V(g_q2)))))
plot(g_q2, vertex.size = vertex.size, edge.width = E(g_q2)$MI*edge.scalar, cex = text.size, layout = coords_q2, vertex.shape = "none")
```

```{r}
dat_q3 = dat_seq %>%
  filter(C0 >= cutoffs_50_75[1] & C0 <= cutoffs_50_75[2]) %>%
  select(all_of(ps1), all_of(ps2), C0)

d_q3 = as.data.frame(matrix(nrow = 1225, ncol = 3))
colnames(d_q3) = c("Var1", "Var2", "MI")
z=1
for(i in 1:49) {
  for(j in (i+1):50) {
    d_q3[z, "Var1"] = i
    d_q3[z, "Var2"] = j
    d_q3[z, "MI"] = mutinformation(dat_q3[,paste0("X", i, "mono")], dat_q3[,paste0("X", j, "mono")])
    z=z+1
  }
}

d_q3 = d_q3 %>%
  arrange(desc(MI))
g_q3 = graph_from_data_frame(d_q3[1:number.vertices,], directed=FALSE)
coords_q3 <- layout_in_circle(g_q3, order =
                                order(as.numeric(names(V(g_q3)))))
plot(g_q3, vertex.size = vertex.size, edge.width = E(g_q3)$MI*edge.scalar, cex = text.size, layout = coords_q3, vertex.shape = "none")
```

```{r}
dat_q4 = dat_seq %>%
  filter(C0 >= cutoffs_75_100[1] & C0 <= cutoffs_75_100[2]) %>%
  select(all_of(ps1), all_of(ps2), C0)

d_q4 = as.data.frame(matrix(nrow = 1225, ncol = 3))
colnames(d_q4) = c("Var1", "Var2", "MI")
z=1
for(i in 1:49) {
  for(j in (i+1):50) {
    d_q4[z, "Var1"] = i
    d_q4[z, "Var2"] = j
    d_q4[z, "MI"] = mutinformation(dat_q4[,paste0("X", i, "mono")], dat_q4[,paste0("X", j, "mono")])
    z=z+1
  }
}

d_q4.ord <- d_q4

vertex.size = 5
edge.scalar = 800
text.size = 7
number.vertices = 150

d_q4 = d_q4 %>%
  arrange(desc(MI))
g_q4 = graph_from_data_frame(d_q4[1:number.vertices,], directed=FALSE)
coords_q4 <- layout_in_circle(g_q4, order =
                                order(as.numeric(names(V(g_q4)))))
plot(g_q4, vertex.size = vertex.size, edge.width = E(g_q4)$MI*edge.scalar, cex = text.size, layout = coords_q4, vertex.shape = "none")
```

```{r}
library(gridExtra)
```

```{r}
vertex.size = 5
edge.scalar = 400
text.size = 10
number.vertices = 150
par(mfrow = c(2,1))
plot(g_q1, vertex.size = vertex.size, edge.width = E(g_q1)$MI*edge.scalar, cex = text.size, layout = coords_q1, vertex.shape = "none", main="Nucleotides, First Quartile")
# plot(g_q2, vertex.size = vertex.size, edge.width = E(g_q2)$MI*edge.scalar, cex = text.size, layout = coords_q2, vertex.shape = "none", main="Nucleotides, Second Quartile")
# plot(g_q3, vertex.size = vertex.size, edge.width = E(g_q3)$MI*edge.scalar, cex = text.size, layout = coords_q3, vertex.shape = "none", main="Nucleotides, Third Quartile")
plot(g_q4, vertex.size = vertex.size, edge.width = E(g_q4)$MI*edge.scalar, cex = text.size, layout = coords_q4, vertex.shape = "none", main="Nucleotides, Fourth Quartile")


mono_int_indices = c(6,9,10,12,15,21)
d_q1.nc = d_q1 %>%
  filter(Var2 - Var1 >= 3) %>%
  mutate(d = Var2 - Var1) %>%
  mutate(color_edge = (d %in% mono_int_indices))
g_q1.nc = graph_from_data_frame(d_q1.nc[1:number.vertices,], directed=FALSE)
coords_q1.nc <- layout_in_circle(g_q1.nc, order =
                                   order(as.numeric(names(V(g_q1.nc)))))
E(g_q1.nc)$color[E(g_q1.nc)$color_edge == TRUE] <- 'red'
E(g_q1.nc)$color[E(g_q1.nc)$color_edge == FALSE] <- 'grey'
plot(g_q1.nc, vertex.size = vertex.size, edge.width = E(g_q1.nc)$MI*edge.scalar, cex = text.size, layout = coords_q1.nc, vertex.shape = "none")
d_q4.nc = d_q4 %>%
  filter(Var2 - Var1 >= 3) %>%
  mutate(d = Var2 - Var1) %>%
  mutate(color_edge = (d %in% mono_int_indices))
g_q4.nc = graph_from_data_frame(d_q4.nc[1:number.vertices,], directed=FALSE)
coords_q4.nc <- layout_in_circle(g_q4.nc, order =
                                   order(as.numeric(names(V(g_q4.nc)))))
E(g_q4.nc)$color[E(g_q4.nc)$color_edge == TRUE] <- 'red'
E(g_q4.nc)$color[E(g_q4.nc)$color_edge == FALSE] <- 'grey'
plot(g_q4.nc, vertex.size = vertex.size, edge.width = E(g_q4.nc)$MI*edge.scalar, cex = text.size, layout = coords_q4.nc, vertex.shape = "none")
# save.image(file = "MI_nucleotide_noncodon")

d_q2.nc = d_q2 %>%
  filter(Var2 - Var1 >= 3)
d_q3.nc = d_q3 %>%
  filter(Var2 - Var1 >= 3)

d_q1.nc.display = d_q1.nc %>%
  mutate(d = Var2 - Var1) %>%
  top_n(20, MI)

d_q2.nc.display = d_q2.nc %>%
  mutate(d = Var2 - Var1) %>%
  top_n(20, MI)

d_q3.nc.display = d_q3.nc %>%
  mutate(d = Var2 - Var1) %>%
  top_n(20, MI)

d_q4.nc.display = d_q4.nc %>%
  mutate(d = Var2 - Var1) %>%
  top_n(20, MI)

### OR

d_q1.nc.display = d_q1.nc %>%
  mutate(d = Var2 - Var1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q2.nc.display = d_q2.nc %>%
  mutate(d = Var2 - Var1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q3.nc.display = d_q3.nc %>%
  mutate(d = Var2 - Var1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q4.nc.display = d_q4.nc %>%
  mutate(d = Var2 - Var1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

grid.table(d_q1.nc.display)
grid.table(d_q2.nc.display)
grid.table(d_q3.nc.display)
grid.table(d_q4.nc.display)
```


```{r}
# Nucleotide, Non-Codon Regions
vertex.size = 5
edge.scalar = 400
text.size = 10
number.vertices = 100
par(mfrow = c(2,1))
d_q1.nc = d_q1 %>%
  filter(Var2 - Var1 >= 3)
g_q1.nc = graph_from_data_frame(d_q1.nc[1:number.vertices,], directed=FALSE)
coords_q1.nc <- layout_in_circle(g_q1.nc, order =
                                   order(as.numeric(names(V(g_q1.nc)))))
plot(g_q1.nc, vertex.size = vertex.size, edge.width = E(g_q1.nc)$MI*edge.scalar, cex = text.size, layout = coords_q1.nc, vertex.shape = "none", main="First Quartile")
d_q4.nc = d_q4 %>%
  filter(Var2 - Var1 >= 3)
g_q4.nc = graph_from_data_frame(d_q4.nc[1:number.vertices,], directed=FALSE)
coords_q4.nc <- layout_in_circle(g_q4.nc, order =
                                   order(as.numeric(names(V(g_q4.nc)))))
plot(g_q4.nc, vertex.size = vertex.size, edge.width = E(g_q4.nc)$MI*edge.scalar, cex = text.size, layout = coords_q4.nc, vertex.shape = "none", main="Fourth Quartile")
#Saved image as "MI_nucleotide_noncodon_q4.png"
```




```{r}
d_q1.ps2 = as.data.frame(matrix(nrow = 1176, ncol = 3))
colnames(d_q1.ps2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:48) {
  for(j in (i+1):49) {
    d_q1.ps2[z, "Var1"] = i
    d_q1.ps2[z, "Var2"] = j
    d_q1.ps2[z, "MI"] = mutinformation(dat_q1[,paste0("X", i, "di")], dat_q1[,paste0("X", j, "di")])
    z=z+1
  }
}

vertex.size = 5
edge.scalar = 2
text.size = 10
number.vertices = 150
d_q1.ps2 = d_q1.ps2 %>%
  arrange(desc(MI))

# g_q1.ps2 = graph_from_data_frame(d_q1.ps2[1:number.vertices,], directed=FALSE)

# plot(g_q1.ps2, vertex.size = vertex.size, edge.width = E(g_q1.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q1.ps2, vertex.shape = "none")
```

```{r}
d_q2.ps2 = as.data.frame(matrix(nrow = 1176, ncol = 3))
colnames(d_q2.ps2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:48) {
  for(j in (i+1):49) {
    d_q2.ps2[z, "Var1"] = i
    d_q2.ps2[z, "Var2"] = j
    d_q2.ps2[z, "MI"] = mutinformation(dat_q2[,paste0("X", i, "di")], dat_q2[,paste0("X", j, "di")])
    z=z+1
  }
}

d_q2.ps2 = d_q2.ps2 %>%
  arrange(desc(MI))
# g_q2.ps2 = graph_from_data_frame(d_q2.ps2[1:number.vertices,], directed=FALSE)
# plot(g_q2.ps2, vertex.size = vertex.size, edge.width = E(g_q2.ps2)$MI*edge.scalar, cex = text.size)
```

```{r}
d_q3.ps2 = as.data.frame(matrix(nrow = 1176, ncol = 3))
colnames(d_q3.ps2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:48) {
  for(j in (i+1):49) {
    d_q3.ps2[z, "Var1"] = i
    d_q3.ps2[z, "Var2"] = j
    d_q3.ps2[z, "MI"] = mutinformation(dat_q3[,paste0("X", i, "di")], dat_q3[,paste0("X", j, "di")])
    z=z+1
  }
}

d_q3.ps2 = d_q3.ps2 %>%
  arrange(desc(MI))
# g_q3.ps2 = graph_from_data_frame(d_q.ps2[1:number.vertices,], directed=FALSE)

# plot(g_q3.ps2, vertex.size = vertex.size, edge.width = E(g_q3.ps2)$MI*edge.scalar, cex = text.size)
```

```{r}
d_q4.ps2 = as.data.frame(matrix(nrow = 1176, ncol = 3))
colnames(d_q4.ps2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:48) {
  for(j in (i+1):49) {
    d_q4.ps2[z, "Var1"] = i
    d_q4.ps2[z, "Var2"] = j
    d_q4.ps2[z, "MI"] = mutinformation(dat_q4[,paste0("X", i, "di")], dat_q4[,paste0("X", j, "di")])
    z=z+1
  }
}

d_q4.ps2 = d_q4.ps2 %>%
  arrange(desc(MI))
# plot(g_q4.ps2, vertex.size = vertex.size, edge.width = E(g_q4.ps2)$MI*edge.scalar, cex = text.size)
```

```{r}
vertex.size = 15
edge.scalar = 20
text.size = 5
number.vertices = 200

par(mfrow = c(2,2))

g_q1.ps2 = graph_from_data_frame(d_q1.ps2[49:(48+number.vertices),], directed=FALSE)
coords_q1.ps2 <- layout_in_circle(g_q1.ps2, order =
                                    order(as.numeric(names(V(g_q1.ps2)))))
plot(g_q1.ps2, vertex.size = vertex.size, edge.width = E(g_q1.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q1.ps2, vertex.shape = "none", main="Dinucleotides, First Quartile")


g_q2.ps2 = graph_from_data_frame(d_q2.ps2[49:(48+number.vertices),], directed=FALSE)
coords_q2.ps2 <- layout_in_circle(g_q2.ps2, order =
                                    order(as.numeric(names(V(g_q2.ps2)))))
plot(g_q2.ps2, vertex.size = vertex.size, edge.width = E(g_q2.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q2.ps2, vertex.shape = "none", main="Dinucleotides, Second Quartile")


g_q3.ps2 = graph_from_data_frame(d_q3.ps2[49:(48+number.vertices),], directed=FALSE)
coords_q3.ps2 <- layout_in_circle(g_q3.ps2, order =
                                    order(as.numeric(names(V(g_q3.ps2)))))
plot(g_q3.ps2, vertex.size = vertex.size, edge.width = E(g_q3.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q3.ps2, vertex.shape = "none", main="Dinucleotides, Third Quartile")


g_q4.ps2 = graph_from_data_frame(d_q4.ps2[49:(48+number.vertices),], directed=FALSE)
coords_q4.ps2 <- layout_in_circle(g_q4.ps2, order =
                                    order(as.numeric(names(V(g_q4.ps2)))))
plot(g_q4.ps2, vertex.size = vertex.size, edge.width = E(g_q4.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q4.ps2, vertex.shape = "none", main="Dinucleotides, Fourth Quartile")


vertex.size = 15
edge.scalar = 400
text.size = 5
number.vertices = 100

vertex.size = 5
edge.scalar = 200
text.size = 10
number.vertices = 150

di_int_indices = c(3,4,5,8,9,10,11,14,20)
d_q1.ps2 = d_q1.ps2 %>%
  filter(Var2 - Var1 >= 3) %>%
  mutate(d = Var2 - Var1 - 1) %>%
  mutate(color_edge = (d %in% di_int_indices))
g_q1.ps2 = graph_from_data_frame(d_q1.ps2[1:number.vertices,], directed=FALSE)
coords_q1.ps2 <- layout_in_circle(g_q1.ps2, order =
                                    order(as.numeric(names(V(g_q1.ps2)))))
E(g_q1.ps2)$color[E(g_q1.ps2)$color_edge == TRUE] <- 'red'
E(g_q1.ps2)$color[E(g_q1.ps2)$color_edge == FALSE] <- 'grey'
plot(g_q1.ps2, vertex.size = vertex.size, edge.width = E(g_q1.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q1.ps2, vertex.shape = "none")

d_q2.ps2 = d_q2.ps2 %>%
  filter(Var2 - Var1 >= 3)
g_q2.ps2 = graph_from_data_frame(d_q2.ps2[1:number.vertices,], directed=FALSE)
coords_q2.ps2 <- layout_in_circle(g_q2.ps2, order =
                                    order(as.numeric(names(V(g_q2.ps2)))))
plot(g_q2.ps2, vertex.size = vertex.size, edge.width = E(g_q2.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q2.ps2, vertex.shape = "none", main="Dinucleotides, Second Quartile")

d_q3.ps2 = d_q3.ps2 %>%
  filter(Var2 - Var1 >= 3)
g_q3.ps2 = graph_from_data_frame(d_q3.ps2[1:number.vertices,], directed=FALSE)
coords_q3.ps2 <- layout_in_circle(g_q3.ps2, order =
                                    order(as.numeric(names(V(g_q3.ps2)))))
plot(g_q3.ps2, vertex.size = vertex.size, edge.width = E(g_q3.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q3.ps2, vertex.shape = "none", main="Dinucleotides, Third Quartile")

d_q4.ps2 = d_q4.ps2 %>%
  filter(Var2 - Var1 >= 3) %>%
  mutate(d = Var2 - Var1 - 1) %>%
  mutate(color_edge = (d %in% di_int_indices))
g_q4.ps2 = graph_from_data_frame(d_q4.ps2[1:number.vertices,], directed=FALSE)
coords_q4.ps2 <- layout_in_circle(g_q4.ps2, order =
                                    order(as.numeric(names(V(g_q4.ps2)))))
E(g_q4.ps2)$color[E(g_q4.ps2)$color_edge == TRUE] <- 'red'
E(g_q4.ps2)$color[E(g_q4.ps2)$color_edge == FALSE] <- 'grey'
plot(g_q4.ps2, vertex.size = vertex.size, edge.width = E(g_q4.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q4.ps2, vertex.shape = "none")



plot(g_q1.ps2, vertex.size = vertex.size, edge.width = E(g_q1.ps2)$MI*edge.scalar, cex = text.size, layout = coords_q1.ps2, vertex.shape = "none")
#Saved image as "MI_dinucleotide_noncodon_q4.png"

d_q1.ps2.nc.display = d_q1.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  top_n(20, MI)

d_q2.ps2.nc.display = d_q2.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  top_n(20, MI)

d_q3.ps2.nc.display = d_q3.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  top_n(20, MI)

d_q4.ps2.nc.display = d_q4.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  top_n(20, MI)

### OR

d_q1.ps2.nc.display = d_q1.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q2.ps2.nc.display = d_q2.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q3.ps2.nc.display = d_q3.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

d_q4.ps2.nc.display = d_q4.ps2 %>%
  mutate(d = Var2 - Var1 - 1) %>%
  group_by(d) %>%
  summarise(avg_MI = round(mean(MI), 5)) %>%
  top_n(4, avg_MI) %>%
  arrange(desc(avg_MI))

grid.table(d_q1.ps2.nc.display)
grid.table(d_q2.ps2.nc.display)
grid.table(d_q3.ps2.nc.display)
grid.table(d_q4.ps2.nc.display)
```


```{r}
## Nucleotide Heatmaps:
map_q1.nc <- d_q1.nc %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.008416209)) + 
  ylab("Nucleotide Position 2") +
  xlab("Nucleotide Position 1")

map_q2.nc <- d_q2.nc %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.008416209)) + 
  ylab("Nucleotide Position 2") +
  xlab("Nucleotide Position 1")

map_q3.nc <- d_q3.nc %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.008416209)) + 
  ylab("Nucleotide Position 2") +
  xlab("Nucleotide Position 1")

map_q4.nc <- d_q4.nc %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.008416209)) + 
  ylab("Nucleotide Position 2") +
  xlab("Nucleotide Position 1")

map_q1.nc
map_q2.nc
map_q3.nc
map_q4.nc
```

```{r}
## Dinucleotide Heatmaps:
map_q1.ps2 <- d_q1.ps2 %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.02689924)) + 
  ylab("Dinucleotide Position 2") +
  xlab("Dinucleotide Position 1")

map_q2.ps2 <- d_q2.ps2 %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.02689924)) + 
  ylab("Dinucleotide Position 2") +
  xlab("Dinucleotide Position 1")

map_q3.ps2 <- d_q3.ps2 %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.02689924)) + 
  ylab("Dinucleotide Position 2") +
  xlab("Dinucleotide Position 1")

map_q4.ps2 <- d_q4.ps2 %>%
  ggplot(aes(x=Var1, y=Var2, fill=MI)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(0, 0.02689924)) + 
  ylab("Dinucleotide Position 2") +
  xlab("Dinucleotide Position 1")

map_q1.ps2
map_q2.ps2
map_q3.ps2
map_q4.ps2
```


```{r}
# Find marginal distribution of A,T,C,G across entire train set:
marg.A = sum(dat$A)/(nrow(dat)*50)
marg.T = sum(dat$T)/(nrow(dat)*50)
marg.C = sum(dat$C)/(nrow(dat)*50)
marg.G = sum(dat$G)/(nrow(dat)*50)

marg.nc = c(marg.A, marg.T, marg.C, marg.G)

# Simulation Study on mutual information significance
set.seed(50)
# N number of simulations
N = 10000
# M number of "sequences" in each simulation
M = 18533
sim.sequences.nc <- vector("list", N)
# Generate "sequences"
sim.sequences.nc <- lapply(sim.sequences.nc, function(x) {
  matrix(nrow = M, ncol = 2, data = sample(1:4, 2*M, replace = TRUE, prob = marg.nc))
})
# Calculate mutual information for each sequence
sim.MI.nc <- lapply(sim.sequences.nc, function(x) {
  mutinformation(x[,1], x[,2])
})
sim.MI.nc <- unlist(sim.MI.nc)

d_q1.nc.sim <- d_q1.nc %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.nc)/N
  }))

d_q2.nc.sim <- d_q2.nc %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.nc)/N
  }))

d_q3.nc.sim <- d_q3.nc %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.nc)/N
  }))

d_q4.nc.sim <- d_q4.nc %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.nc)/N
  }))

ggplot() +
  geom_density(data=data.frame(Simulated = sim.MI.nc), aes(x = Simulated, color = "Simulated")) + 
  geom_density(data = d_q1.nc.sim, aes(x = MI, color = "Q1")) + 
  geom_density(data = d_q2.nc.sim, aes(x = MI, color = "Q2")) +
  geom_density(data = d_q3.nc.sim, aes(x = MI, color = "Q3")) + 
  geom_density(data = d_q4.nc.sim, aes(x = MI, color = "Q4"))

## Total number of edges in each quartile: 1128
nrow(d_q1.nc.sim)

# Number of significant (alpha < .05) edges in each quartile:
sum(d_q1.nc.sim$pval < .05) # 959
sum(d_q2.nc.sim$pval < .05) # 743
sum(d_q3.nc.sim$pval < .05) # 713
sum(d_q4.nc.sim$pval < .05) # 894

sum(d_q1.nc.sim$pval == 0) # 484
sum(d_q2.nc.sim$pval == 0) # 313
sum(d_q3.nc.sim$pval == 0) # 264
sum(d_q4.nc.sim$pval == 0) # 402

table(d_q1.nc.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1) %>%
        select(d))
table(d_q2.nc.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1) %>%
        select(d))
table(d_q3.nc.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1) %>%
        select(d))
table(d_q4.nc.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1) %>%
        select(d))
```

```{r}
# Find marginal distribution of dinucleotides across entire train set:
marg.AA = sum(dat$AA)/(nrow(dat)*49)
marg.AT = sum(dat$AT)/(nrow(dat)*49)
marg.AC = sum(dat$AC)/(nrow(dat)*49)
marg.AG = sum(dat$AG)/(nrow(dat)*49)
marg.TA = sum(dat$TA)/(nrow(dat)*49)
marg.TT = sum(dat$TT)/(nrow(dat)*49)
marg.TC = sum(dat$TC)/(nrow(dat)*49)
marg.TG = sum(dat$TG)/(nrow(dat)*49)
marg.CA = sum(dat$CA)/(nrow(dat)*49)
marg.CT = sum(dat$CT)/(nrow(dat)*49)
marg.CC = sum(dat$CC)/(nrow(dat)*49)
marg.CG = sum(dat$CG)/(nrow(dat)*49)
marg.GA = sum(dat$GA)/(nrow(dat)*49)
marg.GT = sum(dat$GT)/(nrow(dat)*49)
marg.GC = sum(dat$GC)/(nrow(dat)*49)
marg.GG = sum(dat$GG)/(nrow(dat)*49)

marg.ps2 = c(marg.AA, marg.AC, marg.AG, marg.AT,
             marg.CA, marg.CC, marg.CG, marg.CT,
             marg.GA, marg.GC, marg.GG, marg.GT,
             marg.TA, marg.TC, marg.TG, marg.TT)

set.seed(50)
# N number of simulations
N = 10000
# M number of "sequences" in each simulation
M = 18533
sim.sequences.ps2 <- vector("list", N)
# Generate "sequences"
sim.sequences.ps2 <- lapply(sim.sequences.ps2, function(x) {
  matrix(nrow = M, ncol = 2, data = sample(1:16, 2*M, replace = TRUE, prob = marg.ps2))
})
# Calculate mutual information for each sequence
sim.MI.ps2 <- lapply(sim.sequences.ps2, function(x) {
  mutinformation(x[,1], x[,2])
})
sim.MI.ps2 <- unlist(sim.MI.ps2)

d_q1.ps2.sim <- d_q1.ps2 %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.ps2)/N
  }))

d_q2.ps2.sim <- d_q2.ps2 %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.ps2)/N
  }))

d_q3.ps2.sim <- d_q3.ps2 %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.ps2)/N
  }))

d_q4.ps2.sim <- d_q4.ps2 %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.ps2)/N
  }))

ggplot() +
  geom_density(data=data.frame(Simulated = sim.MI.ps2), aes(x = Simulated, color = "Simulated")) + 
  geom_density(data = d_q1.ps2.sim, aes(x = MI, color = "Q1")) + 
  geom_density(data = d_q2.ps2.sim, aes(x = MI, color = "Q2")) +
  geom_density(data = d_q3.ps2.sim, aes(x = MI, color = "Q3")) + 
  geom_density(data = d_q4.ps2.sim, aes(x = MI, color = "Q4"))

## Total number of edges in each quartile: 1081
nrow(d_q1.ps2.sim)

# Number of significant (alpha < .05) edges in each quartile:
sum(d_q1.ps2.sim$pval < .05) # 1066
sum(d_q2.ps2.sim$pval < .05) # 1013
sum(d_q3.ps2.sim$pval < .05) # 953
sum(d_q4.ps2.sim$pval < .05) # 1025

sum(d_q1.ps2.sim$pval == 0) # 902
sum(d_q2.ps2.sim$pval == 0) # 669
sum(d_q3.ps2.sim$pval == 0) # 517
sum(d_q4.ps2.sim$pval == 0) # 772

table(d_q1.ps2.sim %>%
        filter(pval==0) %>%
        select(d))
table(d_q2.ps2.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1 - 1) %>%
        select(d))
table(d_q3.ps2.sim %>%
        filter(pval==0) %>%
        mutate(d = Var2 - Var1 - 1) %>%
        select(d))
table(d_q4.ps2.sim %>%
        filter(pval==0) %>%
        select(d))
```

## TO DO: 

Apply to different libraries, see if similar structures hold - Random library should show even more pronounced structures since there is no codon dependency

Try removing connections within three nucleotides (codon regions)

Try finding the average mutual information betweeen dinucleotides, then simulate to find the p-values associated with each mutual information score - what is the probability of finding such a high/such a low mutual information? 

Could also 


7/5:

Do 0-1 classification for includes A,T vs not includes A,T - recreate heatmaps and rerun simulations.

7/11:

Complete 0-1 classification

Perform simulation on entire training dataset - find MI on entire training dataset and find significant edges using this simulation.

Find a way to simulate within these quartile-based datasets in a way that makes sense - the vast majority of edges should not be significant
  - Simulate entire dataset and then (somehow) divide into quartiles?
  
7/18:

Perform simulation analysis on random library - we expect .05 of pairs to be significant

Is there a multiple comparisons issue?

Perform simulation based on marginal distributions of trinucleotides - actually build simulated 50bp sequences based on these marginal distributions. Then find mutual informations for each pair? 
  
Elastic net for feature selection

Come up with a feature that captures the ~10bp periodicity.

Increase the efficiency of feature selection in lightGBM or try other models - lasso or something similar.
  
```{r}
# Find mutual information across entire train dataset:
# (as opposed to only within quartiles)

d_full = as.data.frame(matrix(nrow = 1225, ncol = 3))
colnames(d_full) = c("Var1", "Var2", "MI")
z=1
for(i in 1:49) {
  for(j in (i+1):50) {
    d_full[z, "Var1"] = i
    d_full[z, "Var2"] = j
    d_full[z, "MI"] = mutinformation(dat_seq[,paste0("X", i, "mono")], dat_seq[,paste0("X", j, "mono")])
    z=z+1
  }
}

d_full = d_full %>%
  arrange(desc(MI))

d_full.nc = d_full %>%
  filter(Var2 - Var1 >= 3)

## Dinucleotides:

d_full.ps2 = as.data.frame(matrix(nrow = 1176, ncol = 3))
colnames(d_full.ps2) = c("Var1", "Var2", "MI")
z=1
for(i in 1:48) {
  for(j in (i+1):49) {
    d_full.ps2[z, "Var1"] = i
    d_full.ps2[z, "Var2"] = j
    d_full.ps2[z, "MI"] = mutinformation(dat_seq[,paste0("X", i, "di")], dat_seq[,paste0("X", j, "di")])
    z=z+1
  }
}

d_full.ps2 = d_full.ps2 %>%
  arrange(desc(MI))

d_full.ps2 = d_full.ps2 %>%
  filter(Var2 - Var1 >= 3)
```



```{r}
# Simulation Study on mutual information significance - full dataset
set.seed(50)
# N number of simulations
N = 10000
# M number of "sequences" in each simulation
M = 74131
sim.sequences.nc.full <- vector("list", N)
# Generate "sequences"
sim.sequences.nc.full <- lapply(sim.sequences.nc.full, function(x) {
  matrix(nrow = M, ncol = 2, data = sample(1:4, 2*M, replace = TRUE, prob = marg.nc))
})
# Calculate mutual information for each sequence
sim.MI.nc.full <- lapply(sim.sequences.nc.full, function(x) {
  mutinformation(x[,1], x[,2])
})
sim.MI.nc.full <- unlist(sim.MI.nc.full)

d_full.nc.sim <- d_full.nc %>%
  mutate(pval = sapply(MI, function(x) {
    sum(x < sim.MI.nc.full)/N
  }))

ggplot() +
  geom_density(data=data.frame(Simulated = sim.MI.nc.full), aes(x = Simulated)) + 
  geom_histogram(data = d_full.nc.sim, aes(x = MI, fill = "Full dataset", alpha = .01))

## Total number of edges: 1128
nrow(d_full.nc.sim)

# Number of significant (alpha < .05) edges in each quartile:
sum(d_full.nc.sim$pval < .05) # 1093

sum(d_full.nc.sim$pval < .01) # 1034

plot(d_full.nc.sim$MI)
abline(h=.0025)

sum(d_full.nc.sim$MI > .0025) # 47

d_full.nc.sim.temp = d_full.nc.sim %>%
  filter(MI > .0025)
```


```{r}
# # Log ratio of MI for q4 to q1:
# 
# log.rat.ps1 <- log(d_q4.ord$MI/d_q1.ord$MI)
# 
# d_log.rat.ps1 <- d_q1.ord %>%
#   mutate(LogRatio = log.rat.ps1, 
#          d=as.factor(Var2-Var1)) %>%
#   select(-MI) %>%
#   arrange(desc(LogRatio))
# 
# ##FIXME - need to adjust d_q1.ps2, etc. so that they are in order (i.e., Var1, Var2 go 1:50)
# log.rat.ps2 <- log(d_q4.ps2.ord$MI/d_q1.ps2.ord$MI)
```


```{r}
# d_q1.ps2 = d_q1.ps2 %>%
#   mutate(difference = Var2-Var1) %>%
#   mutate(across(difference, as.factor))
# 
# summary(lm(MI~difference, data = d_q1.ps2))$adj.r.squared
# 
# plot(d_q1.ps2$difference, d_q4.ps2$MI)
# 
# 
# 
# d_q2.ps2 = d_q2.ps2 %>%
#   mutate(difference = Var2-Var1) %>%
#   mutate(across(difference, as.factor))
# 
# summary(lm(MI~difference, data = d_q2.ps2))$adj.r.squared
# 
# plot(d_q2.ps2$difference, d_q4.ps2$MI)
# 
# 
# 
# d_q3.ps2 = d_q3.ps2 %>%
#   mutate(difference = Var2-Var1) %>%
#   mutate(across(difference, as.factor))
# 
# summary(lm(MI~difference, data = d_q3.ps2))$adj.r.squared
# 
# plot(d_q3.ps2$difference, d_q4.ps2$MI)
# 
# 
# 
# d_q4.ps2 = d_q4.ps2 %>%
#   mutate(difference = Var2-Var1) %>%
#   mutate(across(difference, as.factor))
# 
# summary(lm(MI~difference, data = d_q4.ps2))$adj.r.squared
# 
# plot(d_q4.ps2$difference, d_q4.ps2$MI)
```











